---
title: "R-Basics"
author: "karthik"
date: '2022-07-05'
output:
  word_document: default
  html_document: default
---
# Array

In R, Arrays can contain multi-dimensional rectangular shaped data storage structure. “Rectangular” in the sense, each row is having the same length and similarly for each column and other dimensions. Matrices are a special type of two — dimensional arrays.

Arrays can contain only homogeneous data, i.e. elements having similar data type. In R, an array can be created in using the array() function. Arrays take vectors in the form of input and use the values in the dim parameter for creating an array.

### Creating an Array
- An array in R can be created with the use of array() function. 
- List of elements is passed to the array() functions along with the dimensions as required.

- Syntax:

- ```array(data, dim = (nrow, ncol, nmat), dimnames=names)```
- nrow : Number of rows 
- ncol : Number of columns 
- nmat : Number of matrices of dimensions nrow * ncol
- dimnames : Default value = NULL.

### Uni-Dimensional Array
- A vector is a uni-dimensional , which is specified by a single dimension, length.
- A Vector can be created using "c()" function. A list of values is passed to the c() function to create a vector.
```{r}
#Single Dimension Array
a.vec<-c(1:9) 
a.vec # vector
arr.1d<-array(a.vec)  # creating array
arr.1d #array
is.array(arr.1d)
is.vector(arr.1d) # arr.1d is not a vector
```
## Multi-Dimensional Array
- A two-dimensional matrix is an array specified by a fixed number of rows and columns, each containing the same data type.
- A matrix is created by using array() function to which the values and the dimensions are passed
```{r}
# arranges data from 2 to 5 # in two matrices of dimensions 5x3
arr = array(2:5, dim = c(5, 3, 2))
print(arr)
```
```{r}
#access the 3D Array 
#arr[row,col,matrix]
arr[c(1:3),c(1:2),c(1)]



```
```{r}
#create array from constant LETTERS  #4D data
print(array(LETTERS,dim=c(3,3,2,2)))
```
### Naming of Arrays
- The row names, column names and matrices names are specified as a vector of the number of rows, number of columns and number of matrices respectively. By default, the rows, columns and matrices are named by their index values.
```{r}
#Naming of array
row_names <- c("row1", "row2")   # row name vector
col_names <- c("col1", "col2", "col3")  #col name vector
mat_names <- c("Mat1", "Mat2") # matrix name vector
arr = array(2:14, dim = c(2, 3, 2),dimnames = list(row_names,col_names, mat_names))
print (arr)
```
### Access the array using label names
```{r}
# Access the array using label names
arr["row1","col2","Mat1"]
arr[ ,"col2","Mat1"]
arr[,,"Mat1"]

```
### Access the array using label numeric index
```{r}
arr[c(-1),c(1:2),1:2]
```
### Access the array using label numeric and label  index
```{r}
arr[c(-1),c(1:3),c("Mat1","Mat2")]
arr['row1','col1',]
arr['row1','col1','Mat1']
```
# subset to Matrix from array
```{r}
mat1.from.array<-arr[,,1]
mat1.from.array
class(mat1.from.array)
is.matrix((mat1.from.array))
```
# Calculations Across Array Elements
- We can do calculations across the elements in an array using the apply() function.
- Syntax: apply(x, margin, fun)
Following is the description of the parameters used −
    x is an array.
    margin is the name of the data set used.
    fun is the function to be applied across the elements of the array.
```{r}
# Create two vectors of different lengths.
vector1 <- c(20:23)
vector2 <- c(10,11,12,13,14,15)
# Take these vectors as input to the array.
new.array <- array(c(vector1,vector2),dim = c(3,3,2))
print(new.array)
# Use apply to calculate the sum of the rows across all the matrices.
result <- apply(new.array,c(1), sum) # row Sum of all matrix
print(result)    

result2 <- apply(new.array,c(2), sum) # col Sum of all matrix
print(result2)    

result2 <- apply(new.array,c(1,2), sum) # row and col Sum of all matrix
print(result2)
```
### 5. Factors
- Categorical (nominal) and ordered categorical (ordinal) variables in R are called factors.
- Factors are the r-objects which are created using a vector. 
- It stores the vector along with the distinct values of the elements in the vector as labels. The labels are always character irrespective of whether it is numeric or character or Boolean etc. in the input vector. They are useful in statistical modeling.

### Creation of factors
- Factors are created using the factor() function. 
- factor(x = character(), levels, labels = levels, exclude = NA, ordered = is.ordered(x), nmax = NA)
- The nlevels functions gives the count of levels.

### Attributes of Factors in R Language
    x: It is the vector that needs to be converted into a factor.
    Levels: It is a set of distinct values which are given to the input vector x.
    Labels: It is a character vector corresponding to the number of labels.
    Exclude: This will mention all the values you want to exclude.
    Ordered: This logical attribute decides whether the levels are ordered.
    nmax: It will decide the upper limit for the maximum number of levels.


```{r}
# Create a vector.
apple_colors <- c('green','green','yellow','red','red','red','green') #character vector
#find unique element in vector
unique(apple_colors)
#find duplicate elements in vector
dup.log.vec<-duplicated(apple_colors)
apple_colors[dup.log.vec]
# converting vector into factor
factor.apple<-factor(apple_colors)
factor.apple
# check obj is factor or not?
is.factor(factor.apple)
is.ordered(factor.apple)
# order factor
z <- factor(LETTERS[3:1], ordered = TRUE)
z
is.ordered(z)
order(z)
#level of factor object
levels(z)
levels(factor.apple)
#exclude 
x <- factor(c(1, 2, NA,NA,3,5,5,2,2,1,1,1), exclude = NA)
x
#nmax
x <- factor(c(1, 2, NA,NA,3,5,5,2,2,1,1,1), exclude = NA,nmax =5)
x
#number of levels in factor object
nlevels(x)
```






















